---
title: "Mitochondrial Imputaion QC"
author: "Shea Andrews"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: true
    number_sections: false
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
editor_options: 
  chunk_output_type: console
params:
  wgs.map: wgs.map
  wgs.ped: wgs.ped
  wgs.vcf: wgs.vcf
  typ.map: typ.map
  typ.ped: typ.ped
  typ.vcf: typ.vcf
  imp.map: imp.map
  imp.ped: imp.ped
  imp.vcf: imp.vcf
  imp.info: imp.info
  info.cut: info.cut
  rwd: rwd
---

```{r setup, echo=FALSE, warning =FALSE, message=FALSE}
knitr::opts_knit$set(root.dir = params$rwd)

suppressPackageStartupMessages(library(tidyverse))
library(ggforce)
library(readxl)
library(HiMC); data(nodes)

##  Function
#source('scripts/import_snps.R', chdir = FALSE)

##  Function for calculating mathews correlation coefficent (MCC)
# Moddifed from the mccr package
mccr <- function (act, pred) 
{
  TP <- sum(act %in% 1 & pred %in% 1)
  TN <- sum(act %in% 0 & pred %in% 0)
  FP <- sum(act %in% 0 & pred %in% 1)
  FN <- sum(act %in% 1 & pred %in% 0)
  denom <- as.double(TP + FP) * (TP + FN) * (TN + FP) * (TN + FN)
  if (any((TP + FP) == 0, (TP + FN) == 0, (TN + FP) == 0, (TN + FN) == 0)) 
    denom <- 1
  mcc <- ((TP * TN) - (FP * FN))/sqrt(denom)
  return(mcc)
}

```
## Human610-Quadv1_B-b37 

```{r readin, echo=FALSE, warning =FALSE, message=FALSE}
##  Read in Thousand Genomes WGS 
#     Plink .ped files 
#     VCF Files
wgs_1kg.ped <- generate_snp_data(params$wgs.map, params$wgs.ped)
wgs_1kg.vcf <- read_tsv(params$wgs.vcf, comment = '##', na = c(".", "", "NA"))

##  Read in Thousand Genomes WGS - Typed only
#     Plink .ped files 
typ_1kg.ped <- generate_snp_data(params$typ.map, params$typ.ped)
typ_1kg.vcf <- read_tsv(params$typ.vcf, comment = '##', na = c(".", "", "NA"))

##  Read in Thousand Genomes WGS - Imputed only
#     Plink .ped files 
imp_1kg.ped <- generate_snp_data(params$imp.map, params$imp.ped) 
imp_1kg.ped <- imp_1kg.ped[,-c(grep("\\<189\\>", colnames(imp_1kg.ped)), grep("\\<16183\\>", colnames(imp_1kg.ped)))]
imp_1kg.vcf <- read_tsv(params$imp.vcf, comment = '##', na = c(".", "", "NA"))
imp_1kg.info <- read_delim(params$imp.info, delim = " ")

```

## Data Munging 
```{r, echo=FALSE, warning =FALSE, message=FALSE}

##  obtain intersect of SNPs from imputed and wgs 
snp.intersect <- wgs_1kg.vcf %>% 
  select(POS) %>% 
  semi_join(imp_1kg.vcf, by = 'POS') %>%
  mutate(POS = paste0('mt', POS))

##  Munge imputed dataframe
# Drop all columns but position and genotypes
imp_1kg <- imp_1kg.vcf %>% 
  select(-`#CHROM`, -ID, -REF, -ALT, -QUAL, -FILTER, -INFO, -FORMAT) %>% 
  distinct(POS, .keep_all = T) %>% 
  mutate(POS = paste0('mt', POS))

# Transpose dataframe
imp_1kg <- imp_1kg %>%
  gather(key = var_name, value = value, 2:ncol(imp_1kg)) %>% 
  spread_(key = names(imp_1kg)[1],value = 'value')

# substitute allels for NA, 0, 1 calls, change formate to interger
imp_1kg <- as.tibble(sapply(imp_1kg, function(x) gsub('\\:.*', "", x)))
imp_1kg[imp_1kg == './.'] <- NA
imp_1kg[imp_1kg == '0/0'] <- 0
imp_1kg[imp_1kg == '1/1'] <- 1
imp_1kg <- imp_1kg %>% 
  select(var_name, snp.intersect$POS) %>% 
  mutate_at(vars(contains('mt')), as.integer)

```

```{r, echo=FALSE, warning =FALSE, message=FALSE}
##  Munge ADNI wgs dataframe
# Drop all columns but position and genotypes
wgs_1kg <- wgs_1kg.vcf %>% 
  select(-`#CHROM`, -ID, -REF, -ALT, -QUAL, -FILTER, -INFO, -FORMAT) %>% 
  distinct(POS, .keep_all = T) %>% 
  mutate(POS = paste0('mt', POS))

# Transpose dataframe
wgs_1kg <- wgs_1kg %>%
  gather(key = var_name, value = value, 2:ncol(wgs_1kg)) %>% 
  spread_(key = names(wgs_1kg)[1],value = 'value')

# substitute allels for NA, 0, 1 calls, change formate to interger
wgs_1kg[wgs_1kg == './.'] <- NA
wgs_1kg[wgs_1kg == '0/0'] <- 0
wgs_1kg[wgs_1kg == '1/1'] <- 1
wgs_1kg <- wgs_1kg %>% 
  select(var_name, snp.intersect$POS) %>% 
  mutate_at(vars(contains('mt')), as.integer) 


```

```{r, echo=FALSE, warning =FALSE, message=FALSE}
imp_1kg.info <- mutate(imp_1kg.info, info_comb = ifelse(info_type0 == -1, info,info_type0 ))
imp_1kg.info <- mutate(imp_1kg.info, himc = ifelse(position %in% c(825, 1018, 1438, 1719, 1736, 2092, 3505, 3552, 3594, 4580, 4769, 4917, 4977, 5178, 5442, 6371, 7028, 8251, 8414, 8468, 8703, 9042, 9055, 9347, 9950, 10115, 10398, 10398, 10400, 10550, 11177, 11251, 11947, 12007, 12308, 12705, 13263, 13368, 13506, 13708, 13789, 14178, 14318, 14470, 14560, 14668, 14766, 15043, 15326, 15452, 15535, 16111, 16189, 16391), 'yes', 'no'))
```

## SNP Concordance Statisitics
Compare the genotype calls of the imputed SNPs to those of the SNPs from the WGS dataset.
```{r, echo=FALSE}
## calculate MCC between imputed and typed SNPs
mccr.geno <- unlist(map2(wgs_1kg[,2:ncol(wgs_1kg)], imp_1kg[,2:ncol(imp_1kg)], function(a,b) mccr(a,b)))

##  Calculate concordance between imputed and typed SNPs
concodance.geno <- unlist(map2(imp_1kg[,2:ncol(imp_1kg)], wgs_1kg[,2:ncol(wgs_1kg)], function(a, b) sum(a == b, na.rm = T)/length(a)))

##  Calculate allele frequency of typed SNPs from WGS data
af <- unlist(map(wgs_1kg[,2:ncol(wgs_1kg)], function(a) sum(a, na.rm = T)/length(a)))

##  Dataframe for summary stats
summary.stats <- tibble(mtSNP = colnames(wgs_1kg[,2:ncol(wgs_1kg)]), 
                        pos = as.integer(gsub('mt', '', colnames(wgs_1kg[,2:ncol(wgs_1kg)]))),
                        af = af,
                        mcc = mccr.geno,
                        concodance = concodance.geno)

##  merge on info.score file
summary.stats <- left_join(summary.stats, select(imp_1kg.info, c(-snp_id, -rs_id)), by = c('pos' = 'position')) 
summary.stats <- mutate(summary.stats, info.cat = cut_width(summary.stats$info, 0.25, boundary = 0))
```


```{r, echo=T}
##  basic summary stats
summary.stats %>% count(af > 0.01)
summary.stats %>% count(mcc > 0.4)
summary.stats %>% count(concodance < 0.9)
summary.stats %>% count(info > 0.3); summary.stats %>% count(info > 0.5)
```

###MCC 
[Mathews Correlation Coefficent](https://en.wikipedia.org/wiki/Matthews_correlation_coefficient) (MCC), which takes into account true and false positives and negatives and is generally regarded as a balanced measure which can be used even if the classes are of very different sizes.
```{r, echo=F, message = FALSE, warning=F, fig.width=7.5, fig.height=4.7, fig.align='center', fig.cap= 'Fig. 1: MCC of Imputed SNPs'}
##  Plots by bp position - all SNPs
ggplot(summary.stats, aes(x = pos, y = mcc, colour = info.cat, size = af, shape = himc)) + geom_point() + theme_bw() + 
  labs(x = 'mtDNA position', y = 'MCC', title = 'Imputed mtSNP allele vs sequenced mtSNP allele') + 
  guides(colour=guide_legend(title="Impute2 Info")) + 
  guides(size=guide_legend(title="Allele Frequency"))
```

###Concordance
[Concordance](https://en.wikipedia.org/wiki/Concordance_(genetics)#Genotyping_studies) is a measure of the percentage of SNPs that are measured as identical.
```{r, echo=F, message = FALSE, warning=F, fig.width=7.5, fig.height=4.7, fig.align='center', fig.cap= 'Fig. 2: Concordance of Imputed SNPs'}
ggplot(summary.stats, aes(x = pos, y = concodance, colour = info.cat, size = af)) + geom_point() + theme_bw() + 
  labs(x = 'mtDNA position', y = 'Concordance', title = 'Imputed mtSNP allele vs sequenced mtSNP allele') + 
  guides(colour=guide_legend(title="Impute2 Info")) + 
  guides(size=guide_legend(title="Allele Frequency"))
```

###[INFO Score](https://mathgen.stats.ox.ac.uk/impute/impute_v2.html#info_metric_details)
```{r, echo=F, message = FALSE, warning=F, fig.width=7.5, fig.height=4.7, fig.align='center', fig.cap= 'Fig. 3: Info of Imputed SNPs'}

ggplot(summary.stats, aes(x = pos, y = info_comb, colour = as.factor(type), shape = himc, size = af)) + geom_point() + theme_bw() + 
  labs(x = 'mtDNA position', y = 'Info Score', title = 'Info Score of imputed mtSNPs') + 
  guides(colour=guide_legend(title="SNP Type")) 

```

## Haplogroup Concordance 

```{r haplogorups, echo=FALSE, warning =FALSE, message=FALSE}
## Assign haplogroup assignments using Hi-MC
#   WGS 
MTwgs.classifications <- HiMC::getClassifications(wgs_1kg.ped)

#   Typed Only
MTtyp.classifications <- HiMC::getClassifications(typ_1kg.ped)

#   Imputed 
## remove sites with poor info scores (< 0.3)
rm.info <- filter(summary.stats, info > params$info.cut)
imp_1kg.filtered <- imp_1kg.ped[,colnames(imp_1kg.ped) %in% c('Individual', rm.info$pos)]

MTimp.classifications <- HiMC::getClassifications(imp_1kg.filtered)

MT_haps <- MTwgs.classifications %>% 
  left_join(MTtyp.classifications, by = 'Individual', suffix = c("_wgs", "_typ")) %>% 
  left_join(MTimp.classifications, by = 'Individual') %>%
  rename(full_path_imp = full_path, haplogroup_imp = haplogroup) %>%
  as.tibble()
```

### Alluvial Diagram

```{r haplogorups2, echo=FALSE, warning =FALSE, message=FALSE}

## Count pairs of haplogroups of imputed and WGS assignments
hap.match_typ <- MT_haps %>%
  count(haplogroup_typ, haplogroup_wgs) %>% 
  mutate(match = haplogroup_typ == haplogroup_wgs) 

## Use ggforce to tidy data for geom_parallel Sets 
# Requires developmental version of ggforce
dat_ggforce_typ <- hap.match_typ  %>%
  gather_set_data(1:2) %>%       
  arrange(x,haplogroup_wgs,desc(haplogroup_typ))

# Label Names for Aluvial diagram
dat_ggforce_typ$z <- factor(dat_ggforce_typ$y)

dat_ggforce_typ$y <- as.numeric(factor(dat_ggforce_typ$y))

dat_ggforce_typ$y <- str_pad(dat_ggforce_typ$y, 3, pad = "0")

# label Names for Imputed haplogroups
labs1_typ <- dat_ggforce_typ %>% 
  filter(x == 'haplogroup_typ') %>% 
  arrange(y) %>% 
  dplyr::distinct(z)
labs1_typ <- as.character(labs1_typ$z)

# label Names for WGS haplogroups
labs2_typ <- dat_ggforce_typ %>% 
  filter(x == 'haplogroup_wgs') %>% 
  arrange(y) %>% 
  dplyr::distinct(z)
labs2_typ <- as.character(labs2_typ$z)
```

```{r haplogorups3, echo=FALSE, warning =FALSE, message=FALSE, fig.align='center', fig.cap= 'Fig. 4: Info of Imputed SNPs'}
ggplot(dat_ggforce_typ, aes(x = x, id = id, split = y, value = n)) +
  geom_parallel_sets(aes(fill = match), alpha = 0.5, axis.width = 0.2) +
  geom_parallel_sets_labels(colour = 'black', angle = 0, label = c(labs1_typ, labs2_typ), size = 3) + 
  theme_classic() + theme(legend.position = 'bottom') + 
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        text = element_text(size=12)) + 
  scale_x_discrete(labels=c("Typed", "WGS")) + 
  labs(x = "Mitochondrial Haplogroups") + scale_fill_brewer(palette = 'Set1')

#ggsave('/Users/sheaandrews/Dropbox/Research/PostDoc-MSSM/3_mitoWAX/Drafts/Figures/mtBeforeImp.pdf', units = 'in', height = 7.5, width = 7.5)
```

```{r haplogorups4, echo=FALSE, warning =FALSE, message=FALSE}
## Count pairs of haplogroups of imputed and WGS assignments
hap.match_imp <- MT_haps %>%
  count(haplogroup_imp, haplogroup_wgs) %>% 
  mutate(match = haplogroup_imp == haplogroup_wgs) 

## Use ggforce to tidy data for geom_parallel Sets 
# Requires developmental version of ggforce
dat_ggforce_imp <- hap.match_imp  %>%
  gather_set_data(1:2) %>%       
  arrange(x,haplogroup_wgs,desc(haplogroup_imp))

# Label Names for Aluvial diagram
dat_ggforce_imp$z <- factor(dat_ggforce_imp$y)

dat_ggforce_imp$y <- as.numeric(factor(dat_ggforce_imp$y))

dat_ggforce_imp$y <- str_pad(dat_ggforce_imp$y, 3, pad = "0")

# label Names for Imputed haplogroups
labs1_imp <- dat_ggforce_imp %>% 
  filter(x == 'haplogroup_imp') %>% 
  arrange(y) %>% 
  dplyr::distinct(z)
labs1_imp <- as.character(labs1_imp$z)

# label Names for WGS haplogroups
labs2_imp <- dat_ggforce_imp %>% 
  filter(x == 'haplogroup_wgs') %>% 
  arrange(y) %>% 
  dplyr::distinct(z)
labs2_imp <- as.character(labs2_imp$z)

```

```{r haplogorups5, echo=FALSE, warning =FALSE, message=FALSE}
ggplot(dat_ggforce_imp, aes(x = x, id = id, split = y, value = n)) +
  geom_parallel_sets(aes(fill = match), alpha = 0.5, axis.width = 0.2) +
  geom_parallel_sets_labels(colour = 'black', angle = 0, label = c(labs1_imp, labs2_imp), size = 3) + 
  theme_classic() + theme(legend.position = 'bottom') + 
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        text = element_text(size=12)) + 
  scale_x_discrete(labels=c("Imputed", "WGS")) + 
  labs(x = "Mitochondrial Haplogroups") + scale_fill_brewer(palette = 'Set1')

#ggsave('/Users/sheaandrews/Dropbox/Research/PostDoc-MSSM/3_mitoWAX/Drafts/Figures/mtAfterImp.pdf', units = 'in', height = 7.5, width = 7.5)
```







































